---
title: "Cybersecurity Workforce Personality Analysis"
author: "Matthew DeLoia, Data Scientist, Defense Systems"
date: "Columbus, Georgia"
output: 
  ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#load libraries
library(Rmisc)
library(foreign)
library (memisc)
library(tidyverse)
library(readxl)
library(eRm)
library(ggcorrplot)
library(corrplot)
library(chron)
library(hms)
library(mclust)
library(imputeMissings)
library(randomForest)
library(plotly)
library(ggdendro)
library(cluster)
library(wordcloud)
library(wordcloud2)
library(stringr)
library(tm)
library(gganimate)
library(pracma)
library(lubridate)

likertNum <- function(x){
        case_when(
                x == "Very Accurate" ~ 5,
                x == "Moderately Accurate" ~ 4,
                x == "Neither Accurate Nor Inaccurate" ~ 3,
                x == "Moderately Inaccurate" ~ 2,
                x == "Very Inaccurate" ~ 1
        )
}

likertNum2 <- function(x){
        case_when(
                x == "VERY ACCURATE" ~ 5,
                x == "MODERATELY ACCURATE" ~ 4,
                x == "NEITHER INACC NOR ACCUR" ~ 3,
                x == "MODERATELY INACCURATE" ~ 2,
                x == "VERY INACCURATE" ~ 1
        )
}

#load Cyber Summit Questionnnaire data
df <- read_excel("CyberSummitQuestionnaire_Data11Aug20.xls")
df <- df %>% rownames_to_column("ID") 
n_CyberPro <- df %>% filter(CyberPro=="Yes") %>% count()
df$Duration <- as.hms(df$Duration)

cogntive <- 88:94
df[cogntive] <- lapply(df[cogntive], as.factor) 
#Impute the mode
df[4:10] <- impute( df[4:10], object = NULL, method = "median/mode", flag = FALSE)
df[88:94] <- impute( df[88:94], object = NULL, method = "median/mode", flag = FALSE)

#Load and Process Eugene-Springfield dataset
df_ESDemo <- as.data.set(spss.portable.file("demographics.por"), to.data.frame=T)
df_ESDemo <-as.data.frame(df_ESDemo)
df_ES <- as.data.set(spss.portable.file("IPIP2539-1.por"), to.data.frame=T)
df_ES <- as.data.frame(df_ES)
df_ES2 <- df_ES %>% dplyr::select (ID, A1:A99, C1:C9, M1:M9) %>% gather(A1:A99, C1:C9, M1:M9, key=Question, value = Response, na.rm = TRUE)
df_ES2$Response <- toupper(df_ES2$Response)
df_ES3 <- df_ES2 %>% 
        mutate_at(vars(`Response`), likertNum2) %>% 
        spread(key=Question, value=Response) 
df_ES4 <- df_ES %>%  dplyr::select(-(A1:A99),-(C1:C9), -(M1:M9)) %>% left_join(df_ES3, by="ID")
df_ES5 <- df_ESDemo %>% left_join(df_ES4, by="ID")
df_ES5$ID <- as.character(df_ES5$ID)
df_ESscored <- df_ES5 %>% group_by(ID) %>% 
        mutate(Anxiety = (sum(H926,H980, S3, H761, X107 ) ),      
               Social_Astuteness = (sum(H1137, P432, P451, V96, V153)),
               Breadth_of_Interest = (sum(X211, V140, V2, V330, H1265) ), 
               Complexity = (sum(H1276, Q228, H1267, H1284, V256) ), 
               Social_Confidence = (sum(H1110,S29,P364, H1366, H16 ) ), 
               Cooperativeness = (sum(P420,E147,D107,E68, H974)),  
               Empathy = (sum(D19, X253,H173,	D6,Q256) ), 
               Energy_Level = (sum(H54,	H555,H554,Q151,H79) ), 
               Innovation = (sum(H53,	V259,V226, H447,  X225) ),  
               Organization  = (sum(V98,Q163,V261, H258, V12) ), 
               Responsibility = (sum(H152,V303, D10, E80,V170) ), 
               Risk_Taking = (sum(S19,S4, R77, R58, Q126) ), 
               Sociability = (sum(D70,X161, H78, H985, Q238) ), 
               Tolerance = (sum(H148, Q59, X59, D113, X90) ),
               Traditional_Values = (sum(D88, V46, Q44, H296, X76) ) ) %>%
  dplyr::select (ID, SEX, AGE, EMPLOY, Anxiety:Traditional_Values) %>%
  dplyr::rename(Age = AGE, Gender = SEX) %>% na.omit() %>%
  mutate(Group = "WorkPopulation")
df_ESscored$Gender <- gsub("M", "Male", df_ESscored$Gender )
df_ESscored$Gender <- gsub("F", "Female", df_ESscored$Gender )

#filter to desired comparative population
df_ESscored2 <- df_ESscored %>% filter(EMPLOY %in% c("full time", "part time", "retired"), Age>=22 & Age<=75, Gender=="Male") %>%  dplyr::select(Group, ID, Gender, Age, Anxiety:Traditional_Values )  %>% na.omit()
summaryES <- df_ESscored2 %>% gather(Anxiety:Traditional_Values, key=Dimension, value=Score) %>% summarySE(groupvars = "Dimension", measurevar = "Score" )

anxiety <- summaryES %>% filter(Dimension=="Anxiety")
breadth_of_interest <- summaryES %>% filter(Dimension=="Breadth_of_Interest")
complexity <- summaryES %>% filter(Dimension=="Complexity")
cooperativeness <- summaryES %>% filter(Dimension=="Cooperativeness")
empathy <- summaryES %>% filter(Dimension=="Empathy")
energy_level <- summaryES %>% filter(Dimension=="Energy_Level")
innovation <- summaryES %>% filter(Dimension=="Innovation")
organization <- summaryES %>% filter(Dimension=="Organization")
responsibility <- summaryES %>% filter(Dimension=="Responsibility")
risk_taking <- summaryES %>% filter(Dimension=="Risk_Taking")
sociability <- summaryES %>% filter(Dimension=="Sociability")
social_astuteness <- summaryES %>% filter(Dimension=="Social_Astuteness")
social_confidence <- summaryES %>% filter(Dimension=="Social_Confidence")
tolerance <- summaryES %>% filter(Dimension=="Tolerance")
traditional_values <- summaryES %>% filter(Dimension=="Traditional_Values")
responsibility <- summaryES %>% filter(Dimension=="Responsibility")

#Processing Demograpic data (Education)
df$FieldofEducation <- gsub("Yes", 1, df$FieldofEducation)
df$FieldofEducation2 <- gsub("Yes", 1, df$FieldofEducation2)
df$FieldofEducation3 <- gsub("Yes", 1, df$FieldofEducation3)
df$FieldofEducation <- gsub("No", 0, df$FieldofEducation)
df$FieldofEducation2 <- gsub("No", 0, df$FieldofEducation2)
df$FieldofEducation3 <- gsub("No", 0, df$FieldofEducation3)
df_ed <- df %>% filter(CyberPro =="Yes") %>% select(ID, Education, FieldofEducation, FieldofEducation2, FieldofEducation3) %>% na.omit()
df_ed$FieldofEducation <- as.numeric(df_ed$FieldofEducation)
df_ed$FieldofEducation2 <- as.numeric(df_ed$FieldofEducation2)
df_ed$FieldofEducation3 <- as.numeric(df_ed$FieldofEducation3)
df_ed2 <- df_ed %>%  group_by(ID) %>%
  mutate(Degrees=if_else(Education=="Bachelors", 1, if_else(Education=="Masters", 2, if_else(Education=="PhD", 3, 0)))) %>% 
  mutate(CS_Degree= sum(FieldofEducation, FieldofEducation2, FieldofEducation3))
cs_degree <- sum(df_ed2$CS_Degree)
degree <- sum(df_ed2$Degrees)
cs_percent <- round(cs_degree/degree, 2)*100

#Processing Demograpic data (Experience)
df_Years <- df %>% filter(CyberPro=="Yes") %>%  
  select(ID, Age, YearsExperience)  %>%  
  gather(Age:YearsExperience, key=Category, value=Years) 
medianAge <- df_Years %>% filter(Category=="Age") 
medianAge <- median(medianAge$Years)
medianExperience <- df_Years %>% filter(Category=="YearsExperience")
medianExperience <- median(medianExperience$Years)

#Processing Demograpic data (Certifications)
df_cert <-df %>% filter(CyberPro=="Yes") %>%  select(Certifications) 
df_cert$Certifications<- gsub("\\(", "", df_cert$Certifications)
df_cert$Certifications<- gsub("\\)", "", df_cert$Certifications)
df_cert2<-df_cert %>% separate(Certifications, c("Certifications1", "Certifications2", "Certifications3", "Certifications4", "Certifications5", "Certifications6", "Certifications7", "Certifications8", "Certifications9"), sep=",") %>%
  gather(Certifications1:Certifications9, key=Category, value=Certifications) %>%
  select (-Category ) %>%  
  na.omit(df_cert)
df_cert3 <- df_cert2 %>% group_by(Certifications) %>% summarise(n=n()/n_CyberPro$n)

#Processing Demograpic data (Hobbies)
df_hobby <-df %>% filter(CyberPro=="Yes") %>%  select(Hobbies) 
df_hobby$Hobbies<- gsub("\\(", "", df_hobby$Hobbies)
df_hobby$Hobbies<- gsub("\\)", "", df_hobby$Hobbies)
df_hobby2<-df_hobby %>% separate(Hobbies, c("Hobbies1", "Hobbies2", "Hobbies3", "Hobbies4", "Hobbies5", "Hobbies6", "Hobbies7", "Hobbies8"), sep=",") %>%
  gather(Hobbies1:Hobbies8, key=Category, value=Hobbies) %>% select (-Category ) %>%
  na.omit(df_hobby) 

#Scoring respondant data        
df_scores <- df %>%   
        mutate_at(vars(P_AN5:P_RE4), likertNum) %>% 
        mutate(Pattern_Q58 = if_else(Pattern_Q58==4, 1, 0),
        Pattern_Q1 = if_else(Pattern_Q1==4, 1, 0),
        `3D_Q24` = if_else(`3D_Q24`==3, 1, 0),
        Verbal_Q4 = if_else(Verbal_Q4==4, 1, 0),
        Verbal_Q16 = if_else(Verbal_Q16==4, 1, 0),
        Matrix_Q55 = if_else(Matrix_Q55==4, 1, 0) )

#Rasch scoring cognitive questions
df_cognitive <- df_scores %>% select(ID, Pattern_Q58:Matrix_Q55) %>% column_to_rownames("ID") 
rm.res <- RM(df_cognitive)
pp <- person.parameter(rm.res)
df_cognitive2 <-  pp$theta.table %>% 
        rownames_to_column("ID") %>% 
        mutate(Proficiency=round(`Person Parameter`, 2)) %>%  select(ID, Proficiency)

#Summary scores for personality and cognitive areas
df_scored <- df_scores %>% dplyr::group_by(ID) %>% 
  mutate(Group="NG_Cyber") %>%
  mutate(Anxiety = (sum(P_AN1, P_AN2, P_AN3, P_AN4, P_AN5 ) ) ,
                Social_Astuteness = (sum(P_AS1, P_AS2, P_AS3, P_AS4, P_AS5	)),
                Breadth_of_Interest = (sum(P_BR1, P_BR2, P_BR3, P_BR4, P_BR5	) ) , 
                Complexity = (sum(P_CM1, P_CM2, P_CM3, P_CM4, P_CM5) ) ,
                Social_Confidence = (sum(P_CD1, P_CD2, P_CD3, P_CD4, P_CD5) ) ,
                Cooperativeness = (sum(P_CO1, P_CO2, P_CO3, P_CO4, P_CO5)),
                Empathy = (sum(P_EM1, P_EM2, P_EM3, P_EM4, P_EM5) ) , 
                Energy_Level = (sum(P_EN1, P_EN2, P_EN3, P_EN4, P_EN5)) ,
                Innovation = (sum(P_IN1, P_IN2, P_IN3, P_IN4, P_IN5) ) , 
                Organization = (sum(P_OR1, P_OR2, P_OR3, P_OR4, P_OR5 ) ) , 
                Responsibility = (sum(P_RE1, P_RE2, P_RE3, P_RE4, P_RE5) ) ,
                Risk_Taking = (sum(P_RI1, P_RI2, P_RI3, P_RI4, P_RI5	) ) , 
                Sociability = (sum(P_SO1, P_SO2, P_SO3, P_SO4, P_SO5 ) ) , 
                Tolerance = (sum(P_TO1, P_TO2, P_TO3, P_TO4, P_TO5 ) ) ,
                Traditional_Values= (sum(P_TR1, P_TR2, P_TR3, P_TR4, P_TR5) )) %>%
  mutate (Cognitive_Score = (sum(Pattern_Q58, Pattern_Q1, `3D_Q24`, Verbal_Q4, Verbal_Q16, Matrix_Q55)) ) %>% 
  left_join(df_cognitive2) %>%  
  select(ID:Education,Group:Cognitive_Score, Proficiency, Duration, SelfEfficacy, Email) 

#Scale socres to ES sample (n=151)
df_scored2 <- df_scored %>%
  group_by(ID) %>%
        mutate(Anxiety= (Anxiety-anxiety$Score) / anxiety$sd)   %>% 
        mutate(Breadth_of_Interest= (Breadth_of_Interest-breadth_of_interest$Score)/breadth_of_interest$sd)  %>% 
        mutate(Complexity= (Complexity - complexity$Score)/ complexity$sd)  %>% 
        mutate(Cooperativeness= (Cooperativeness-cooperativeness$Score)/cooperativeness$sd)  %>% 
        mutate(Empathy= (Empathy-empathy$Score)/empathy$sd) %>% 
        mutate(Energy_Level= (Energy_Level-energy_level$Score)/energy_level$sd) %>% 
        mutate(Innovation= (Innovation-innovation$Score)/innovation$sd)  %>%
        mutate(Organization= (Organization-organization$Score)/organization$sd) %>% 
        mutate(Responsibility= (Responsibility-responsibility$Score)/responsibility$sd) %>% 
        mutate(Risk_Taking= (Risk_Taking-risk_taking$Score)/risk_taking$sd)  %>% 
        mutate(Sociability= (Sociability-sociability$Score)/sociability$sd) %>%
        mutate(Social_Astuteness= (Social_Astuteness-social_astuteness$Score)/social_astuteness$sd) %>%
        mutate(Social_Confidence= (Social_Confidence-social_confidence$Score)/social_confidence$sd)  %>%
        mutate(Tolerance= (Tolerance-tolerance$Score)/tolerance$sd) %>%
        mutate(Traditional_Values= (Traditional_Values-traditional_values$Score)/traditional_values$sd)

#Latent Profile Analysis
lpa_test <- df_scores %>% filter(CyberPro=="Yes") %>% select(P_AN5:P_RE4, ID) %>% column_to_rownames ("ID") 
#clustering <- lpa_test %>%  mutate_all(list(scale))
#BIC <- mclustBIC(clustering)
#plot(BIC)
#summary(BIC)
#mod1 <- Mclust(clustering, modelNames = "VII", G = 1, x = BIC)
#mod1$call
#cluster <- mod1$classification %>% as.factor()
set.seed(98)
kmeans_test <- df_scores %>% filter(CyberPro=="Yes") %>% select(P_AN5:P_RE4, ID) %>% column_to_rownames ("ID")
clustering <- kmeans_test %>% mutate_all(list(scale))
myclusters <- kmeans(clustering, centers = 2, nstart = 25)
cluster <- myclusters$cluster

df_scored2 <- as.data.frame(df_scored2)
df_scored3 <- df_scored2 %>% filter(CyberPro=="Yes") %>%  cbind(cluster)

#visualization of LPA
df_cluster <- df_scored3 %>% gather(Anxiety:Traditional_Values, key=Dimension, value=Score)
df_cluster$cluster<-as.factor(df_cluster$cluster)
df_cluster2 <- df_cluster %>% summarySE (groupvars=c("Dimension", "cluster"), measurevar="Score")%>% 
  mutate(Comparison = if_else(Score<=-.5, "Low", if_else(Score>=.5, "High", "Average")))

df_cluster2$Comparison <- factor (df_cluster2$Comparison, levels = c("Low", "Average", "High"))
```

## Introduction {.build}
Problem Statement:  To understand if highly specialized roles, like cybersecurity, attract workers with similar personality profiles.

Hypothesis to test: Cybersecurity workers are similar in personality to others in the general workforce population.

## Session Aims {.pagenumbers}

- Data Collection and Reporting: report the most prevalent personality profile among cybersecurity professionals attending CyberSummit20

- Analysis & Insights: identify unique personality traits of cybersecurity professionals

- Conclusions: share ideas and opinions through the chat module (so I can benefit from your experience and insight)

## Background (1 of 3)|Cybersecurity Talent Assessment Design

- Demographic Information (Age, Gender, Years Experience, Education, Certifications, Hobbies )
- Personality Assessment (75 Questions, 15 traits)
- Cognitive Assessment (6 Questions: Pattern Matching, Matrix, Verbal Reasoning, 3D Shapes)

*Comparison Group: approximatley 400 individuals with similar demographic profiles from the Eugene-Springfield Community Sample (1993 - 2003)

## Background (2 of 3) | Personality Assessment

This 75 question personality assessment aims to measure 15 underlying personality traits using similar measurements to the <a href="https://w8y7a2j5.stackpathcdn.com/wp-content/uploads/2015/03/JPI-R-Scales.pdf">Jackson Personality Inventory - Revised (JPI-R)</a>.  

The JPI-R measures were selected due to the range of personality traits measured and inclusion of 7 measurements for "analytical" and "dependable," traits previously reported in those performing highly technical work roles.  This set of 15 measures has been useful in vocational counseling.  

This ***JPI-R proxy*** consists of questions drawn from the public domain repository of personality questions  (https://ipip.ori.org/). <br>

## Background (3 of 3)
<div class="columns-2">
***Analytical Scales***

- Complexity 
- Breadth of Interest
- Innovation
- Tolerance

***Extroverted Scales***

- Sociability 
- Social Confidence 
- Energy Level 

***Emotional*** 

- Empathy 
- Anxiety 
- Cooperativeness 

***Opportunistic*** 

- Risk Taking 
- Social Astuteness 

***Dependable***

- Organization 
- Traditional Values 
- Responsibility
</div>

<div class="notes">

- Complexity: seeks intricate solutions to problems
- Tolerance: accepts people even though they may differ from oneself; welcomes dissent
- Cooperativeness: is susceptible to social influence and group pressures; follows suit
- Social Astuteness: skillful at persuading others to achieve a particlular goal; occasionally seen as manipulative of others
- Traditional Values: values traditional customs and beliefs; others may view individual as "old-fashioned"
- Responsibility: strong obligation to be honest and upright; an inflexibly conscience

</div>

## Review of Participant Demographics

- Highest Level of Education
- Age
- Experience
- Certifications
- Hobbies

## Highest Level of Education (n=`r n_CyberPro`)
`r cs_percent`%  post-secondary degrees in computer science or related fields.
```{r Demographics, echo=FALSE, message=FALSE}
df_ed <-df %>% filter (CyberPro=="Yes") %>% select(Education) %>%   group_by(Education) %>% summarise(n=n()/n_CyberPro$n)
df_ed$Education <-  factor(df_ed$Education, levels=c("Secondary Education", "Associates", "Bachelors", "Masters", "PhD"))
df_ed %>% ggplot(aes(x=Education, y=n)) + 
        geom_col(fill="skyblue") + 
        ylab("Percentage") + 
        geom_text(aes( x=Education, y=n, label = round(n, 2), ),color="black", vjust= -.1, hjust = .3, size=3.5) +
         theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) + ylim(0,1) + xlab("")
```

## Age and Years of Experience 
Median age is `r medianAge` years;  median years of experience is `r medianExperience`.
```{r Demographics2, echo=FALSE, message=FALSE}
df_Years %>% ggplot(aes(x=Years)) +
  geom_density(fill="skyblue") + 
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ylab("") + xlab ("") +
  facet_grid(.~Category, scales="free_x") 
```

## Certifications
```{r Certifications, echo=FALSE, message=FALSE}
df_cert3 %>% filter (Certifications != "None", n >.03) %>% ggplot() + geom_col(aes(x=reorder(Certifications, n), y=n), fill="skyblue") +
  geom_text(aes( x=Certifications, y=n, label = round(n, 2), ),color="black", vjust= -.1, hjust = .4, size=3.5) +
  coord_flip() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) + xlab("") + ylab("% of cybersecurity professionals with certificate")
```

## Hobbies 
```{r Hobbies, message=FALSE, warning=FALSE, echo=FALSE}
text2 <- df_hobby2$Hobbies# Create a corpus  
docs2 <- Corpus(VectorSource(text2))
docs2 <- docs2 %>%
  #tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs2 <- tm_map(docs2, content_transformer(tolower))
docs2 <- tm_map(docs2, removeWords, stopwords("english"))
dtm2 <- TermDocumentMatrix(docs2) 
matrix2 <- as.matrix(dtm2) 
words2 <- sort(rowSums(matrix2),decreasing=TRUE) 
df_text2 <- data.frame(word = names(words2),freq=words2)
wordcloud2(data=df_text2, size=.4, color='random-dark', backgroundColor = "lightgray" )
```

## Personality Results (1 of 2)
```{r ScaledScores, echo=FALSE, message=FALSE, warning=FALSE}
personality_results2 <- df_scored2 %>% filter(CyberPro=="Yes") %>% 
  gather(Anxiety:Traditional_Values, key=Dimension, value = Score) %>% group_by(Dimension) %>% 
  arrange(Dimension, Age) %>%  mutate(Average = movavg(Score, 94, type="s"))

personality_results2$Age <- as.integer(personality_results2$Age)
  
q <- personality_results2 %>% filter(Age >=22) %>%     
        ggplot(aes(x=reorder(Dimension, Average, FUN=mean), y=round(Average, 2), color=Dimension)) + 
        geom_point(size=3) +
        coord_flip() + ylim(-1.5,1.5) +
        geom_hline(yintercept=0, linetype="dotted") +
        xlab("") +
     ylab("moving average (z score)") +
        theme(legend.title= element_text(color="black", size=10), legend.position = "blank") 
 
  anim2 <- q + transition_time(Age) +  labs(title = "Age: {frame_time}") + shadow_wake(wake_length = 0.1, alpha = FALSE)
animate(anim2, height=500, width=600)
```

<div class="notes">

Do personality traits modify with Age and Experience?

</div>

## Personality Results (2 of 2) 
```{r ScaledScores2, echo=FALSE, message=FALSE}
personality_results <- df_scored2 %>% filter(CyberPro=="Yes") %>% 
  gather(Anxiety:Traditional_Values, key=Dimension, value = Score) %>%
  summarySE(measurevar = "Score", groupvars = "Dimension") %>%
   mutate(Comparison = if_else(Score<=-.5, "Low", if_else(Score>=.5, "High", "Average")))
personality_results$Comparison <- factor (personality_results$Comparison, levels = c("High", "Average", "Low"))

personality_results %>% 
        mutate(Dimension=reorder(Dimension, Score, FUN=max)) %>%    
        ggplot(aes(x=Dimension, y=round(Score, 2), color=Comparison)) + 
        geom_point() +
        geom_errorbar(aes(ymin=Score-ci, ymax=Score+ci), width=.1 ) +
        coord_flip() + ylim(-2,2) +
        geom_hline(yintercept=0, linetype="dotted") +
        xlab("") +
     ylab("group mean with 95% ci (z score)") +
        scale_color_manual(name="general workforce comparison", values=c("green", "darkgray", "blue")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") + 
  labs(caption = "Note: z scores developed from the E-S Community Sample data to establish norms of the general workforce")
```

<div class="notes">

What aspects of personlity are most stable or like the general workforce?

</div>

## Clustering Analysis

```{r HClustering, echo=FALSE, warning=FALSE}
h_cluster <- hclust(dist(clustering))
h_cluster %>% ggdendrogram(rotate=TRUE, theme_dendro = TRUE, size=1)
```

<div class="notes">

Are there multiple personality profiles within the cybersecurity workforce?

</div>

## Latent Profile Analysis (1 of 3)
cluster 1: n= `r df_scored3 %>% filter(cluster=="1") %>% count()` ; cluster 2: n= `r df_scored3 %>% filter(cluster=="2") %>% count()`
```{r LPAResults, echo=FALSE, warning=FALSE}
df_cluster2 %>% ungroup() %>%
  mutate(Dimension = fct_reorder(Dimension, Score, .fun=max)) %>%
  ggplot(aes (x=Dimension, y=Score, group=cluster)) +
  geom_col(aes(fill=Comparison), size=2) +
  scale_fill_manual(values = c("skyblue", "gray", "lightgreen")) +
  coord_flip() + 
  geom_hline(yintercept = 0, color="red", linetype="dashed") +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
  xlab("") + ylab("cluster mean (z score)") +
  ylim(-1.5,1.5)+
  facet_grid(.~cluster)
```

## Latent Profile Analysis (2 of 3)
```{r LPAResults3, echo=FALSE, message=FALSE, warning=FALSE}
df_cluster3 <- df_cluster2 %>% ungroup() %>%  group_by(Dimension) %>% 
  summarise(clustersd=abs(diff(Score))) %>% 
  right_join(df_cluster)

df_cluster3 %>% ungroup() %>% mutate(Dimension=reorder(Dimension, clustersd, FUN=max)) %>%
  ggplot(aes()) +
    geom_density( aes(x=Score, fill=cluster), alpha=.7 ) +
    ylim(0,.9) +
    geom_hline(yintercept=0, linetype="dotted") +
   scale_linetype_manual(values=c("solid", "dashed")) +
    xlab("") +
   ylab("") +
    scale_fill_manual(values=c("lightgreen", "skyblue")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +facet_wrap(Dimension~.)
```
<div class="notes">

Where are the cluster differences greatest?

</div>

## Latent Profile Analysis (3 of 3)
```{r LPAResults2, echo=FALSE, message=FALSE, warning=FALSE}
df_cluster3 %>% ungroup() %>% summarySE(groupvars=c("Dimension", "cluster", "clustersd"), measurevar="Score") %>% mutate(Dimension=reorder(Dimension, clustersd, FUN=max)) %>% 
        ggplot(aes(x=Dimension, y=round(Score, 2), color=cluster, group=cluster)) + 
        geom_point(size=3) +
        geom_errorbar(aes(ymin=Score-ci, ymax=Score+ci), width=.1, color="darkgray" ) +
        coord_flip() + ylim(-2,2) +
        geom_hline(yintercept=0, linetype="dotted") +
        #geom_line( aes(linetype=cluster) ) +
        #scale_linetype_manual(values=c("blank", "dashed")) +
        xlab("") +
     ylab("cluster mean with 95% ci (z score)") +
        scale_color_manual(values=c("lightgreen", "skyblue")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top")
```
<div class="notes">

Which profile is most distinguished from the general workforce?

</div>

## Comparison of Profiles to Cluster 1 
```{r Comparison, echo=FALSE, message=FALSE, warning=FALSE}
df_selectcluster <- df_scored %>% as.data.frame() %>%  
  filter(CyberPro=="Yes") %>% 
  cbind(cluster) %>% 
  filter(cluster=="1") %>%
  gather(Anxiety:Traditional_Values, key=Dimension, value=Score) %>% 
  summarySE(groupvars = c("cluster", "Dimension"), measurevar = "Score")

anxiety2 <- df_selectcluster %>% filter(Dimension=="Anxiety")
breadth_of_interest2 <- df_selectcluster %>% filter(Dimension=="Breadth_of_Interest")
complexity2 <- df_selectcluster %>% filter(Dimension=="Complexity")
cooperativeness2 <- df_selectcluster %>% filter(Dimension=="Cooperativeness")
empathy2 <- df_selectcluster %>% filter(Dimension=="Empathy")
energy_level2 <- df_selectcluster %>% filter(Dimension=="Energy_Level")
innovation2 <- df_selectcluster %>% filter(Dimension=="Innovation")
organization2 <- df_selectcluster %>% filter(Dimension=="Organization")
responsibility2 <- df_selectcluster %>% filter(Dimension=="Responsibility")
risk_taking2 <- df_selectcluster %>% filter(Dimension=="Risk_Taking")
sociability2 <- df_selectcluster %>% filter(Dimension=="Sociability")
social_astuteness2 <- df_selectcluster %>% filter(Dimension=="Social_Astuteness")
social_confidence2 <- df_selectcluster %>% filter(Dimension=="Social_Confidence")
tolerance2 <- df_selectcluster %>% filter(Dimension=="Tolerance")
traditional_values2 <- df_selectcluster %>% filter(Dimension=="Traditional_Values")
responsibility2 <- df_selectcluster %>% filter(Dimension=="Responsibility")

df_comparison <- df_scored %>%  
        dplyr::select(Group, CyberPro, ID, Gender, Age, Anxiety:Traditional_Values ) %>%
       group_by(ID) %>% 
        mutate(Anxiety= (Anxiety-anxiety2$Score) / anxiety2$sd)   %>% 
        mutate(Breadth_of_Interest= (Breadth_of_Interest-breadth_of_interest2$Score)/breadth_of_interest2$sd)  %>% 
        mutate(Complexity= (Complexity - complexity2$Score)/ complexity2$sd)  %>% 
        mutate(Cooperativeness= (Cooperativeness-cooperativeness2$Score)/cooperativeness2$sd)  %>% 
        mutate(Empathy= (Empathy-empathy2$Score)/empathy2$sd) %>% 
        mutate(Energy_Level= (Energy_Level-energy_level2$Score)/energy_level2$sd) %>% 
        mutate(Innovation= (Innovation-innovation2$Score)/innovation2$sd)  %>%
        mutate(Organization= (Organization-organization2$Score)/organization2$sd) %>% 
        mutate(Responsibility= (Responsibility-responsibility2$Score)/responsibility2$sd) %>% 
        mutate(Risk_Taking= (Risk_Taking-risk_taking2$Score)/risk_taking2$sd)  %>% 
        mutate(Sociability= (Sociability-sociability2$Score)/sociability2$sd) %>%
        mutate(Social_Astuteness= (Social_Astuteness-social_astuteness2$Score)/social_astuteness2$sd) %>%
        mutate(Social_Confidence= (Social_Confidence-social_confidence2$Score)/social_confidence2$sd)  %>%
        mutate(Tolerance= (Tolerance-tolerance2$Score)/tolerance2$sd) %>%
        mutate(Traditional_Values= (Traditional_Values-traditional_values2$Score)/traditional_values2$sd) %>% 
        gather(Anxiety:Traditional_Values, key=Dimension, value=Score)

p <- df_comparison %>%  filter(ID %in% c("2", "3", "4", "47", "6" ,"33", "9", "20")) %>%  ggplot(aes(x=reorder(Dimension, abs(Score), .fun=max), y=Score, color=CyberPro)) + 
  geom_point(size=3) + 
  coord_flip() + ylim(-4,4) +
  geom_hline(yintercept = 0, linetype="dashed") +
  scale_color_manual(values=c("skyblue", "green")) + 
  xlab("") + ylab("z score") + 
  theme(legend.position = "blank") +
  labs(caption = "Note: z scores developed from selected cluster mean and standard deviation values")

# https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r/
  
anim <- p + transition_states(ID, transition_length = 1, state_length = 1) 
animate(anim, height=500, width=600)
```

<div class="notes">

Please vote: Which color do you think highlights the cybersecurity professionals in this sample?

</div>

## Cognitive Proficiency
```{r Proficiency, echo=FALSE, message=FALSE, warning=FALSE}
df_scored %>% filter (CyberPro == "Yes", ID!="3") %>% 
  ggplot(aes(y=Proficiency, x=Age)) + 
  geom_point() + 
  geom_smooth() + labs(caption = "Note: Rasch Scoring used to determine proficiency (theta) as a measure of cognitive ability; zero represents average")
```
<div class="notes">

Is there a relationship between cognitive ability and age?

</div>

## Correlation Analysis (1 of 2)
```{r Correlation, echo=FALSE, message=FALSE, warning=FALSE}
dfcorrplot <- df_ed2 %>% select(ID, Degrees) %>% left_join(df_scored3)  %>% select(ID, SelfEfficacy, Age, YearsExperience, Degrees, Proficiency) %>% column_to_rownames("ID") %>% as.data.frame()

#dfcorrplot$cluster <- as.numeric(dfcorrplot$cluster)
dfcorrplot$SelfEfficacy <-  as.numeric(dfcorrplot$SelfEfficacy)
#dfcorrplot <- scale(dfcorrplot)

corrplot(cor(dfcorrplot), method="color", order="hclust", type="full", addrect=4, cl.lim=c(-1,1), addCoef.col="black", rect.col="green", diag = FALSE, number.font=1, number.digits = 1, number.cex = .7)
```

<div class="notes">

Is "Self-Efficacy" dependent on age and experience?

</div>

## Correlation Analysis (2 of 2)
```{r Correlation2, echo=FALSE, message=FALSE, warning=FALSE}
dfcorrplot <- df_ed2 %>% select(ID, Degrees) %>% left_join(df_scored3)  %>% select(ID, Anxiety:Traditional_Values) %>% column_to_rownames("ID") %>% as.data.frame()

corrplot(cor(dfcorrplot), method="color", order="hclust", type="full", addrect=5, cl.lim=c(-1,1), addCoef.col="black", rect.col="green", diag = FALSE, number.font=1, number.digits = 1, number.cex = .7)
```

## Conclusions {.build}
<div class="blue">
Hypothesis to test: Cybersecurity workers are similar in personality to others in the general workforce population.
</div>


What is your Conclusion: ***Accept*** the Null Hypothesis -or- ***Reject*** the Null Hypothesis?

<div class="notes">

Please vote: Do you Accept the Null Hypothesis -or- do you Reject the Null Hypothesis?

</div>

Can personality profiling predict one's professional interests and ***latent talent*** to perform specialized work roles?

Individual Reports...Coming Soon.

## References
Douglas N. Jackson, Technical Manual: Jackson Personlaity Inventory - Revised, published by Sigma Assessment Systems, August 1994.

International Personality Item Pool, available at: https://ipip.ori.org/

R Core Team (2019). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria. URL
  https://www.R-project.org/.

Eugene-Springfield Community Sample, Harvard Dataverse, available at: https://dataverse.harvard.edu/dataverse/ESCS-Data 

Full Presentation available at: https://mattdeloia.github.io/CyberSummit20/

## Back-up: Gender Differences
```{r TestComparison, echo=FALSE, message=FALSE, warning=FALSE}
df_ESscored <- as.data.frame(df_ESscored)
df_ESscored %>% mutate_at(vars(Anxiety:Traditional_Values), scale) %>%  gather(Anxiety:Traditional_Values, key=Dimension, value=Score) %>% summarySE(groupvars = c("Gender", "Dimension"), measurevar = "Score") %>% 
      mutate(Dimension=reorder(Dimension, Score, FUN=max)) %>%
      ggplot(aes(x=Dimension, y=Score, group=Gender)) +
      geom_errorbar(aes(ymin=Score-ci, ymax=Score+ci), width=.2 ) +
      geom_point( aes(color=Gender), size=3) +
      coord_flip() + ylim(-2,2) +
      scale_color_manual(values=c("tomato", "lightblue")) +
      ylab("gender mean with 95% ci (z score)") +
            xlab("") +  
      theme(legend.position = "top") + 
      labs(caption = "*gender differences by personality trait based on E-S Community Sample data (n~400)")
```


## Back-Up: Trait Descriptions (1 of 3) 
Analytical <br>

 - Complexity: seeks intricate solutions to problems
 - Breadth of Interest: interested in learing about a diversity of things
 - Innovation: motivated to develop novel solutions to problems; values new ideas
 - Tolerance: accepts people even though they may differ from oneself; welcomes dissent
 
## Back-Up: Trait Descriptions (2 of 3) 
Emotional 

 - Empathy: tends to identify closely with other people and their problems
 - Anxiety: tends to worry over inconsequential matters; apprehensive about the future
 - Cooperativeness: is susceptible to social influence and group pressures; follows suit

Extroversion 

- Sociability: eagerly join a varity of social groups; actively social
- Social Confidence: confident in dealing with others; shows presence in interpersonal situations
- Energy level: is active and spirited; does not tire easily; capable of intense work for long periods

## Back-Up: Trait Descriptions (3 of 3) 
Opportunistic 

- Social Astuteness: skillful at persuading others to achieve a particlular goal; occasionally seen as manipulative of others, but ordinarily diplomatic
- Risk Taking: enjoys gambling and taking a chance; willingly exposes self to situations with uncertain outcomes

Dependable 

- Organization: makes effective use of time; not easily distracted
- Traditional Values: values traditional customs and beliefs; others may view individual as "old-fashioned"
- Responsibility: strong obligation to be honest and upright; an inflexibly conscience   